### 2021-11-19 main.Rmd
Fix

---
title: "Woah cool a proper git paper!"
author: "Ryan, Brown, & Platt"
date: "`r format(Sys.Date() + lubridate::years(20),'%A, %B %d, %Y')`"
output: bookdown::pdf_document2
toc: yes
bibliography: references.bib
csl: aag.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
```

```{r rpois, echo = F}
poiss <- rpois(100, 2)
```

```{r child = "intro.Rmd"}
```

\pagebreak

```{r child = "scatter.Rmd"}
```

See figure \@ref(fig:hist) and figure \@ref(fig:scatter) for more details; also read @andersen2021.

\pagebreak

# References