### 2021-11-19 main.Rmd
Fix

---
title: "Woah cool a proper git paper!"
author: "Ryan, Brown, & Platt"
date: "`r format(Sys.Date() + lubridate::years(20),'%A, %B %d, %Y')`"

# Using bookdown makes some things easier, like the \@ref() calls to figures and tables.
output: bookdown::pdf_document2

# Show table of contents?
toc: yes

bibliography: references.bib

# If we don't use AAG, we'll need to change this
csl: aag.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
```

```{r rpois, echo = F}
poiss <- rpois(100, 2)
```

```{r child = "intro.Rmd"}
```

\pagebreak

```{r child = "scatter.Rmd"}
```

You can call data from the main .Rmd from child files! Neat. Let's use this for mapping etc. -- for testing purposes we should define data in each section .Rmd to make things easier, but for the paper itself it'd be noice to only do that once in main.Rmd

See figure \@ref(fig:hist) and figure \@ref(fig:scatter) for more details; also read @andersen2021.

You can use `\@ref(fig:name-of-code-chunk)` to get the figure/table number, rather than manually entering "figure 2, 3, etc." It just outputs the number of the figure.

You can link to Zotero and type `@andersen2021` etc. to enter a Zotero-compatible in-text citation.

\pagebreak

# References