### 2022-06-14 references.Rmd
Quantum finaglery

---
output: word_document
bibliography: bibliography.bib
csl: seg.csl
# nocite: |

---

[@nutsford2013; @ekkel2017; @collins2020; @sakar2018; @jiang2021; @lachowcyz2013; @wang2021; @verheij2008; @vandillen2012; @wood2017; @wheeler2015 @mears2019; @houlden2019; @mears2019a;  @mccal-hosenfeld2014; @cdc2022; @feng2017; @zhang2017; @beyer2014; @mceachen2015; @shin2020; @johnson2018; @feng2020; @thetrustforpublicland2021;  @openstreetmap2022; @runkle2022; @slawsky2022; @usgs2019; @wood2018;  @browning2018; @hoffimann2017; @rigolon2018; @acs2018; @anselin2008; @legendre1993; @rstudioteam2022; @rcoreteam2022; @beyer2014; @houlden2019;  @wang2021; @shin2020; @johnson2018; @xie2020; @feng2020; @chum2015;  @astell-burt2020; @feng2017; @mahec; @alcock2015; @devries2020; @feng2017;  @gresenz2001; @reiss2013; @howell2008; @padgham2017; @sarkar2018]